### 2014-09-23 ApTranscriptome_TR.Rmd

~~~
# download filtered Trinity assembly, uncompress and move
wget http://johnstantongeddes.org/assets/files/Aphaenogaster_transcriptome.tar
# check md5sum
md5sum Aphaenogaster_transcriptome.tar
# fd6dbb0b3e88e1c797c9e74611b245b2

# uncompress and move
tar -xvf Aphaenogaster_transcriptome.tar
mkdir -p results/
mkdir -p results/trinity-full/
mv Trinity_cap3_uclust.fasta results/trinity-full/.
mv Trinity_cap3_uclust_clean.fasta results/trinity-full/.
~~~

@@

```{r download, echo = TRUE}
# URL for annotation file
annotation.URL <- getURL("http://johnstantongeddes.org/assets/files/ApTranscriptome_AnnotationTable_20140113.txt")
# load 
annotation.file <- read.csv(textConnection(annotation.URL), header = TRUE, sep = "\t", stringsAsFactors = FALSE)
str(annotation.file)

@@

~~~
wget ...
mkdir -p /results/trinity-full/sailfish-expression-Trinity-cap3-uclust
mv .. /results/trinity-full/sailfish-expression-Trinity-cap3-uclust
tar -zxvf ...
~~~

@@

Calculate T~on~ for *High* genes in each species.

@@

@@

```{r rData, echo=FALSE}
# save RData for manuscript file
save(TPM.dt.sub, RxNpval, sigtable, chi1, resp.TPM.dt.sub, Ap.type.table, tt2, tt3, md, qlslope, qlmean, GO.stress.Xsq, GO0006950.responsive, Ap.df, w1, w2, w3, Ap.T_on, Ap.BP.resultParentChild, resp.stress.responsive.count, t.mean, t.slope, t.curvature, t.wiggle, t.varint, t.varbim, T_on.high.ttest, T_on.low.ttest, Ap.gsea.union, Ap.gsea.intersect, Ar.int.A22.bim.gsea, Ar.int.A22.low.gsea, file = "ms.RData")
```

>>>>

~~~
# download filtered Trinity assembly, uncompress and move
wget http://johnstantongeddes.org/assets/files/ApTranscriptome/Aphaenogaster_transcriptome.tar
# check md5sum
md5sum Aphaenogaster_transcriptome.tar
# fd6dbb0b3e88e1c797c9e74611b245b2

# move and extract
mkdir -p results/
mkdir -p results/trinity-full/
mv Aphaenogaster_transcriptome.tar results/trinity-full/.
tar -xvf Aphaenogaster_transcriptome.tar
~~~

@@

```{r download, echo = TRUE}
# URL for annotation file
annotation.URL <- getURL("http://johnstantongeddes.org/assets/files/ApTranscriptome/ApTranscriptome_AnnotationTable_20140113.txt")
# load 
annotation.file <- read.csv(textConnection(annotation.URL), header = TRUE, sep = "\t", stringsAsFactors = FALSE)
str(annotation.file)

@@

~~~
# download gene expression quantification
wget http://johnstantongeddes.org/assets/files/ApTranscriptome/sailfish_quant_20140916.tar
# check md5sum
md5sum sailfish_quant_20140916.tar
# 26102c7ef86cf30b5f8e923640378185

# move and extract
mkdir -p /results/trinity-full/sailfish-expression-Trinity-cap3-uclust
mv sailfish_quant_20140916.tar results/trinity-full/.
tar -xvf sailfish_quant_20140916.tar
~~~

@@

Calculate T~on~ for *High* genes in each species. I use the observed (rather than predicted) values for each species as by nature of fitting a quadratic function, the maximum change tends to be at the extremes (38.5 or 0) for the predicted values.

@@

Repeat analysis for when **Intermediate** genes are turned *off*.

```{r T_int_off}
# transcripts expressed at *Intermediate* temperatures in A22
A22.int.TPM.dt.sub <- resp.TPM.dt.sub.pred[J(A22.int.transcripts, "A22")]
setkey(A22.int.TPM.dt.sub, Transcript)
str(A22.int.TPM.dt.sub)

# make data.frame for results
l5 <- length(unique(A22.int.TPM.dt.sub$Transcript))
A22.int.T_off <- data.frame(Transcript = unique(A22.int.TPM.dt.sub$Transcript), colony = rep("ApVT", length = l5), type = rep("Low", length = l5), T_off_high = NA, T_off_low = NA)

# loop across transcripts, calculating T_off at both high and low temperatures
for(i in unique(A22.int.TPM.dt.sub$Transcript)) {
    subdf <- A22.int.TPM.dt.sub[i]

    subdf.high <- subdf[which(subdf$val >= 21), ]
    T_off_high <- subdf.high[median(which(diff(subdf.high$TPM) == min(diff(subdf.high$TPM)))), val]
    A22.int.T_off[which(A22.int.T_off$Transcript == i), "T_off_high"] <- T_off_high

    subdf.low <- subdf[which(subdf$val <= 21), ]
    T_off_low <- subdf.low[median(which(diff(subdf.low$TPM) == min(diff(subdf.low$TPM)))), val]
    A22.int.T_off[which(A22.int.T_off$Transcript == i), "T_off_low"] <- T_off_low
}

# repeat for Ar
Ar.int.TPM.dt.sub <- resp.TPM.dt.sub.pred[J(Ar.int.transcripts, "Ar")]
setkey(Ar.int.TPM.dt.sub, Transcript)
str(Ar.int.TPM.dt.sub)

# make data.frame for results
l6 <- length(unique(Ar.int.TPM.dt.sub$Transcript))
Ar.int.T_off <- data.frame(Transcript = unique(Ar.int.TPM.dt.sub$Transcript), colony = rep("ApVT", length = l6), type = rep("Low", length = l6), T_off_high = NA, T_off_low = NA)

# loop across transcripts, calculating T_off at both high and low temperatures
for(i in unique(Ar.int.TPM.dt.sub$Transcript)) {
    subdf <- Ar.int.TPM.dt.sub[i]

    subdf.high <- subdf[which(subdf$val >= 21), ]
    T_off_high <- subdf.high[median(which(diff(subdf.high$TPM) == min(diff(subdf.high$TPM)))), val]
    Ar.int.T_off[which(Ar.int.T_off$Transcript == i), "T_off_high"] <- T_off_high

    subdf.low <- subdf[which(subdf$val <= 21), ]
    T_off_low <- subdf.low[median(which(diff(subdf.low$TPM) == min(diff(subdf.low$TPM)))), val]
    Ar.int.T_off[which(Ar.int.T_off$Transcript == i), "T_off_low"] <- T_off_low
}

# compare T_off among species
(T_off.int.high.ttest <- t.test(Ar.int.T_off$T_off_high, A22.int.T_off$T_off_high))

(T_off.int.low.ttest <- t.test(Ar.int.T_off$T_off_low, A22.int.T_off$T_off_low))

```

As with genes turned on at high temperatures, *A. carolinensis* **Intermediate** genes are down-regulated on average at 1.2C higher temperatures than *A. picea* **Intermediate** genes.

@@

```{r rData, echo=FALSE}
# save RData for manuscript file
save(TPM.dt.sub, RxNpval, sigtable, chi1, resp.TPM.dt.sub, Ap.type.table, tt2, tt3, md, qlslope, qlmean, GO.stress.Xsq, GO0006950.responsive, Ap.df, w1, w2, w3, Ap.T_on, Ap.BP.resultParentChild, resp.stress.responsive.count, T_on.high.ttest, T_on.low.ttest, T_off.int.high.ttest, T_off.int.low.ttest, Ap.gsea.union, Ap.gsea.intersect, Ar.int.A22.bim.gsea, Ar.int.A22.low.gsea, file = "ms.RData")
```
